---
title: "Listening to sad music during induced sadness"
author: "Joel Larwood"
output:
  html_document:
    code_folding: "hide"
---

```{r}
knitr::opts_chunk$set(tidy = TRUE)
```

People often report listening to sad music when they are experience sadness. It has been argued that people are likely to listen to songs that make them sad when they are sad, especially in cases when they are high in trait rumination. 

This project evaluates this claim further by having people nominate a song that makes them sad before inducing a sad state and asking them to listen to the nominated sad song. Sadness was measured before and after listening to the sad song with measures taken of the emotion mechanisms that are associated with the song and the likelihood of the song being listened to. 

This notebook provides the analysis of the results. The script for data cleaning can be seen [here](/script)

```{r pkgLoad, message=FALSE, results='hold'}

xfun::pkg_attach2("tidyverse", 
                  "here", 
                  "lmerTest", 
                  "apa", 
                  "emmeans"
                  )

```


## Source Files 
Note: This script will not run unless the code chunk options are changed to `eval = TRUE`


```{r source, eval = FALSE}

source(list.files(here::here("code"), pattern = ".R"))

```

```{r}
data <- read_rds(here::here("data", "RMER_November2018_ProcessedWide.rds"))
```

# Get descriptives of the data that includes the spotify variables and the lyric information

## Internal consistency 

```{r measures}

measures <- list(t1 = select(data, deq_1_1:deq_1_4),
                 t2 = select(data, deq_2_1:deq_2_4),
                 t3 = select(data, deq_3_1:deq_3_4),
                 rumination = select(data, rrq_1:rrq_12), 
                 musebaq = select(data, musebaq_1:musebaq_9)
                 )



for (i in measures){
  print(round(psych::alpha(i)[["total"]][["raw_alpha"]], 2))
}

```

# Mean and SD for self report 
```{r meanSD}

psych::describe(select(data, 
                       age, 
                       musebaq:rumination)) 
```
# Logistic Regression rumination and BRECVEMA


```{r}
brecvema_counts <- data %>% 
  select(BSR:Appraisal) %>% 
  map(table)
  
brecvema_counts
```

## Logistic Regression 

The logistic regression indicates that mechanism occurance is not influenced by rumination. 
```{r}

p_extract <- function(...){
  coef()[2, 4]
}

log_reg_brecvema_p <- data %>% 
  select (BSR:Appraisal) %>%
  map(~glm(.x ~ data$rumination, family = binomial, data = data)) %>% 
  map(summary) %>% 
  map(coef) %>% 
  map_dfr(8) %>% 
  pivot_longer(everything(), 
               names_to = "mechanism", 
               values_to = "p") 

log_reg_brecvema_OR <- data %>% 
  select (BSR:Appraisal) %>%
  map(~glm(.x ~ data$rumination, family = binomial, data = data)) %>% 
  map(summary) %>% 
  map(coef) %>% 
  map_dfr(5) %>% 
  pivot_longer(everything(), 
               names_to = "mechanism", 
               values_to = "int") %>% 
  mutate(OR = exp(int))


left_join(log_reg_brecvema_OR, log_reg_brecvema_p) 

```

Nor is the liklihood of listening to a song during sadness, except for contagion and memory

```{r}


log_reg_reguse_p <- data %>% 
  select (BSR:Appraisal) %>%
  map(~glm(.x ~ data$rumination, family = binomial, data = data)) %>% 
  map(summary) %>% 
  map(coef) %>% 
  map_dfr(8) %>% 
  pivot_longer(everything(), 
               names_to = "mechanism", 
               values_to = "p") 

log_reg_reguse_OR <- data %>% 
  select (BSR:Appraisal) %>%
  map(~glm(.x ~ data$reg.use_1, family = binomial, data = data)) %>% 
  map(summary) %>% 
  map(coef) %>% 
  map_dfr(5) %>% 
  pivot_longer(everything(), 
               names_to = "mechanism", 
               values_to = "int") %>% 
  mutate(OR = exp(int))


left_join(log_reg_reguse_OR, log_reg_reguse_p) 
```




# Manipulation check 

```{r manipulationCheck}
apa::t_apa(t.test(data$PostInduction, data$Baseline,
                  paired = TRUE))
```

The manipulation of sadness was successful 

# Pre to Post listening increase 

```{r PrePostSadness}
apa::t_apa(t.test(data$PostListening, data$PostInduction, 
                  paired = TRUE))
```


Sadness also increased from pre to post Listening 

# Effect of Rumination

First I need to make long data 
```{r MakeLong}

data_long <- data %>% 
  pivot_longer(cols = c("PostInduction", "PostListening"), 
               names_to = "Timepoint", 
               values_to = "Sadness")
```


Here I fit a series of general linear mixed models and compare them 

```{r}
# Model 1 is the hypiothesised model 

mod1 <- lmerTest::lmer(Sadness~Timepoint*rumination + (1|ID), 
                       data = data_long, 
                       REML = FALSE)

# Model 2 I control for th interaction with baseline sadness
mod2 <- lmerTest::lmer(Sadness~
                         Timepoint * rumination + 
                         Timepoint * Baseline + 
                         (1|ID), 
                       data = data_long, 
                       REML = FALSE)
# Model 3 I add the theoretically relvant BRECVEMA mechanisms of Contagion 

mod3 <- lmerTest::lmer(Sadness~
                         Timepoint * rumination + 
                         Timepoint * Baseline + 
                         Timepoint* Contagion + 
                         (1|ID), 
                       data = data_long,
                       REML = FALSE)

# Model 4 I add the marginally significant mechanisms predicted by rumination 
mod4 <- lmerTest::lmer(Sadness~
                         Timepoint * rumination + 
                         Timepoint * Baseline + 
                         Timepoint* Contagion + 
                         Timepoint * Conditioning + 
                         Timepoint * Memory +
                         (1|ID), 
                       data = data_long,
                       REML = FALSE)

# Model 5 I add appraisal

mod5 <- lmerTest::lmer(Sadness~
                         Timepoint * rumination + 
                         Timepoint * Baseline + 
                         Timepoint* Contagion + 
                         Timepoint* Appraisal + 
                         Timepoint * Conditioning + 
                         Timepoint * Memory +
                         (1|ID), 
                       data = data_long,
                       REML = FALSE)

mod6 <- lmerTest::lmer(Sadness~
                         Timepoint * rumination + 
                         Timepoint * Baseline + 
                         Timepoint* Contagion + 
                         Timepoint* Appraisal + 
                         Timepoint * Conditioning + 
                         Timepoint * Memory +
                         Timepoint * rumination * Appraisal + 
                         (1|ID), 
                       data = data_long,
                       REML = FALSE)

AIC(mod1, mod2, mod3, mod4, mod5, mod6)
```

From the AIC it can be seen the `mod5` is the best fit

```{r BestModel}

sjPlot::tab_model(mod1, mod5, 
                  show.est = FALSE, 
                  show.std = TRUE, 
                  show.aic = TRUE)
```

I will now set up reference grids for eah of the models 
```{r}

mean_rum <- mean(data$rumination, na.rm = TRUE)
plus_rum <- mean_rum + sd(data$rumination, na.rm = TRUE)
minus_rum <- mean_rum - sd(data$rumination, na.rm = TRUE)

mean_baseline <- mean(data$Baseline, na.rm = TRUE)
plus_baseline <- mean_rum + sd(data$Baseline, na.rm = TRUE)
minus_baseline <- mean_rum - sd(data$Baseline, na.rm = TRUE)

hyp_grid <- emmeans::ref_grid(mod1, 
                              at = list(rumination = c(minus_rum, mean_rum, plus_rum), 
                                        Timepoint = c("PostInduction", "PostListening")))


```

# Hypothesised Model Plot 

## Rumination
```{r}


rum_simple_plot <- emmeans::emmip(hyp_grid, 
                                  formula = rumination ~ Timepoint, 
                                  CIs = TRUE, 
                                  frequentist = TRUE) +
  ggplot2::scale_x_discrete(labels = c("Post Induction", "Post Listening")) + 
  ggplot2::scale_colour_viridis_d(labels = c("-1 SD", "Mean", "+1 SD"),
                                name = "Rumination \nScore")+
  ggplot2::ylab("Predicted Sadness Score") + 
  ggplot2::xlab("Timepoint") +
  ggplot2::ylim(c(10, 18)) +
  ggplot2::theme_classic(base_size = 20, base_family = "Times New Roman") 


rum_simple_plot 


```


## Johnson Neyman 
As there are two levels to the timepoint factor we can look at the Johnson Neyman interval for rumination to indicate at what point of rumination an increase significantly predicted. 

```{r}

jn_data <- data_long %>% 
  mutate(Timepoint = if_else(Timepoint == "PostInduction", 0, 1))

jn_model <- lmer(Sadness~Timepoint + Timepoint * rumination + (1|ID), 
                 data = jn_data, 
                 REML = FALSE)

jn_rum <- interactions::johnson_neyman(model = jn_model, 
                       pred = "Timepoint",
                       modx = "rumination"
                       )

print(jn_rum$bounds)

jn_rum$plot

```

# Best fiting model 
```{r}


mod5_grid <- emmeans::ref_grid(mod5, 
                                   at = list(rumination = c(minus_rum, mean_rum, plus_rum),
                                             Baseline = c(minus_baseline, mean_baseline, plus_baseline), 
                                             Contagion = c("No", "Yes"), 
                                             Conditioning = c("No", "Yes"), 
                                             Appraisal = c("No", "Yes"), 
                                             Memory = c("No", "Yes")))
                                             

```

## Plots 

Now I want to plot each of the interactions.  

```{r}

rum_plot <- emmeans::emmip(mod5_grid, 
               rumination~Timepoint, 
               CIs = TRUE ) + 
  ggplot2::scale_color_discrete(labels = c("-1 SD", "Mean", "+1 SD"),
                                name = "Rumination \nScore") +
  ggplot2::scale_x_discrete(labels = c("Pre", "Post"))
    

baseline_plot <- emmeans::emmip(mod5_grid, 
               Baseline~Timepoint, 
               CIs = TRUE ) +
  ggplot2::scale_color_discrete(labels = c("-1 SD", "Mean", "+1 SD"),
                                name = "Baseline \nScore") +
  ggplot2::scale_x_discrete(labels = c("Pre", "Post"))

contagion_plot <- emmeans::emmip(mod5_grid, 
               Contagion~Timepoint, 
               CIs = TRUE) +
  ggplot2::scale_x_discrete(labels = c("Pre", "Post"))

conditioing_plot <- emmeans::emmip(mod5_grid, 
               Conditioning ~ Timepoint, 
               CIs = TRUE) +
  ggplot2::scale_x_discrete(labels = c("Pre", "Post"))

appraisal_plot <- emmeans::emmip(mod5_grid, 
               Appraisal~ Timepoint, 
               CIs = TRUE) +
  ggplot2::scale_x_discrete(labels = c("Pre", "Post"))

memory_plot <- emmeans::emmip(mod5_grid, 
               Memory ~ Timepoint, 
               CIs = TRUE) +
  ggplot2::scale_x_discrete(labels = c("Pre", "Post"))

cowplot::plot_grid(rum_plot, 
                   baseline_plot, 
                   contagion_plot,
                   appraisal_plot,
                   conditioing_plot, 
                   memory_plot)

```

We can see dense patterning here so I am going to create a scenario where BRECVEMA models are set in the most or least favourable way. In both cases appraisal has been set to No

```{r}

mod5_best <- emmeans::ref_grid(mod5, 
                                   at = list(rumination = c(minus_rum, mean_rum, plus_rum),
                                            # Baseline = c(plus_baseline), 
                                             Contagion = c("No"), 
                                             Conditioning = c("No"), 
                                            Appraisal = c("No"), 
                                             Memory = c("No")))

mod5_worst <- emmeans::ref_grid(mod5, 
                                   at = list(rumination = c(minus_rum, mean_rum, plus_rum),
                                            # Baseline = c(minus_baseline), 
                                             Contagion = c("Yes"), 
                                             Conditioning = c("Yes"), 
                                            Appraisal = c("No"), 
                                             Memory = c("Yes")))

mod5_contagion <- emmeans::ref_grid(mod5, 
                                   at = list(rumination = c(minus_rum, mean_rum, plus_rum),
                                            # Baseline = c(plus_baseline), 
                                             Contagion = c("Yes"), 
                                             Conditioning = c("No"), 
                                            Appraisal = c("No"), 
                                             Memory = c("No")))

mod5_memory <- mod5_best <- emmeans::ref_grid(mod5, 
                                   at = list(rumination = c(minus_rum, mean_rum, plus_rum),
                                            # Baseline = c(plus_baseline), 
                                             Contagion = c("No"), 
                                             Conditioning = c("No"), 
                                            Appraisal = c("No"), 
                                             Memory = c("Yes")))

plot_best <- emmeans::emmip(mod5_best, 
               rumination ~ Timepoint, 
               CIs = TRUE) +
  ggplot2::scale_x_discrete(labels = c("Pre", "Post")) +
  ggplot2::scale_color_discrete(labels = c("-1 SD", "Mean", "+1 SD"),
                                name = "Rumination \nScore") +
  ggplot2::ggtitle(label = "No inductive mechanisms") + 
  ggplot2::ylim(c(8, 16))

plot_worst <- emmeans::emmip(mod5_worst, 
               rumination ~ Timepoint, 
               CIs = TRUE) +
  ggplot2::scale_x_discrete(labels = c("Pre", "Post")) +
  ggplot2::scale_color_discrete(labels = c("-1 SD", "Mean", "+1 SD"),
                                name = "Rumination \nScore") +
    ggplot2::ggtitle(label = "All inductive mechanisms") + 
  ggplot2::ylim(c(8, 16))

plot_mem <- emmeans::emmip(mod5_memory, rumination ~ Timepoint, 
               CIs = TRUE) +
  ggplot2::scale_x_discrete(labels = c("Pre", "Post")) +
  ggplot2::scale_color_discrete(labels = c("-1 SD", "Mean", "+1 SD"),
                                name = "Rumination \nScore") +
    ggplot2::ggtitle(label = "Memories Only") + 
  ggplot2::ylim(c(8, 16))

plot_con <- emmeans::emmip(mod5_memory, rumination ~ Timepoint, 
               CIs = TRUE) +
  ggplot2::scale_x_discrete(labels = c("Pre", "Post")) +
  ggplot2::scale_color_discrete(labels = c("-1 SD", "Mean", "+1 SD"),
                                name = "Rumination \nScore") +
    ggplot2::ggtitle(label = "Contagion Only") + 
  ggplot2::ylim(c(8, 16))

cowplot::plot_grid(plot_best, 
                   plot_worst, 
                   plot_mem,
                   plot_con)


```


## Get pairwise contrasts 
Here I am getting the contrasts for the plotted scenarios 

```{r}

best_contrast <- emmeans::contrast(mod5_best, 
                  "pairwise") %>% 
  broom::tidy() %>% 
  filter(stringr::str_starts(level1, "PostInduction")) %>% 
  tidyr::separate(col = level1, 
                  into = c("Timepoint", "rumination", "baseline", "Contagion", "Appraisal", "Conditioning", "Memory"),
                  sep = ",", 
                  remove = TRUE) %>% 
  tidyr::separate(col = level2, 
                  into = c("Timepoint2", "rumination2", "baseline2", "Contagion2", "Appraisal2", "Conditioning2", "Memory2"),
                  sep = ",", 
                  remove = TRUE) %>%
  filter(Timepoint == "PostInduction" &
           Timepoint2 == "PostListening" & 
           rumination == rumination2 &
           baseline == baseline2 & 
           Contagion == Contagion2 &
           Appraisal == Appraisal2 & 
           Conditioning == Conditioning2 &
           Memory == Memory2) %>% 
  select(-contains("2")) %>% 
  mutate(scenario = "Best", 
         rumination = if_else(rumination < 3.2, "-1 SD", 
                              if_else(rumination > 4, "+1 SD", 
                              "Mean"))) %>% 
  select(scenario, everything(), -Timepoint, -baseline)

worst_contrast <- emmeans::contrast(mod5_worst, 
                  "pairwise") %>% 
  broom::tidy() %>% 
  filter(stringr::str_starts(level1, "PostInduction")) %>% 
  tidyr::separate(col = level1, 
                  into = c("Timepoint", "rumination", "baseline", "Contagion", "Appraisal", "Conditioning", "Memory"),
                  sep = ",", 
                  remove = TRUE) %>% 
  tidyr::separate(col = level2, 
                  into = c("Timepoint2", "rumination2", "baseline2", "Contagion2", "Appraisal2", "Conditioning2", "Memory2"),
                  sep = ",", 
                  remove = TRUE) %>%
  filter(Timepoint == "PostInduction" &
           Timepoint2 == "PostListening" & 
           rumination == rumination2 &
           baseline == baseline2 & 
           Contagion == Contagion2 &
           Appraisal == Appraisal2 & 
           Conditioning == Conditioning2 &
           Memory == Memory2) %>% 
  select(-contains("2")) %>% 
  mutate(scenario = "Worst", 
         rumination = if_else(rumination < 3.2, "-1 SD", 
                              if_else(rumination > 4, "+1 SD", 
                              "Mean"))) %>% 
  select(scenario, everything(), -Timepoint, -baseline)

contagion_contrast <- emmeans::contrast(mod5_contagion, 
                  "pairwise") %>% 
  broom::tidy() %>% 
  filter(stringr::str_starts(level1, "PostInduction")) %>% 
  tidyr::separate(col = level1, 
                  into = c("Timepoint", "rumination", "baseline", "Contagion", "Appraisal", "Conditioning", "Memory"),
                  sep = ",", 
                  remove = TRUE) %>% 
  tidyr::separate(col = level2, 
                  into = c("Timepoint2", "rumination2", "baseline2", "Contagion2", "Appraisal2", "Conditioning2", "Memory2"),
                  sep = ",", 
                  remove = TRUE) %>%
  filter(Timepoint == "PostInduction" &
           Timepoint2 == "PostListening" & 
           rumination == rumination2 &
           baseline == baseline2 & 
           Contagion == Contagion2 &
           Appraisal == Appraisal2 & 
           Conditioning == Conditioning2 &
           Memory == Memory2) %>% 
  select(-contains("2")) %>% 
  mutate(scenario = "Contagion Only", 
         rumination = if_else(rumination < 3.2, "-1 SD", 
                              if_else(rumination > 4, "+1 SD", 
                              "Mean"))) %>% 
  select(scenario, everything(), -Timepoint, -baseline)

memory_contrast <- emmeans::contrast(mod5_memory, 
                  "pairwise") %>% 
  broom::tidy() %>% 
  filter(stringr::str_starts(level1, "PostInduction")) %>% 
  tidyr::separate(col = level1, 
                  into = c("Timepoint", "rumination", "baseline", "Contagion", "Appraisal", "Conditioning", "Memory"),
                  sep = ",", 
                  remove = TRUE) %>% 
  tidyr::separate(col = level2, 
                  into = c("Timepoint2", "rumination2", "baseline2", "Contagion2", "Appraisal2", "Conditioning2", "Memory2"),
                  sep = ",", 
                  remove = TRUE) %>%
  filter(Timepoint == "PostInduction" &
           Timepoint2 == "PostListening" & 
           rumination == rumination2 &
           baseline == baseline2 & 
           Contagion == Contagion2 &
           Appraisal == Appraisal2 & 
           Conditioning == Conditioning2 &
           Memory == Memory2) %>% 
  select(-contains("2")) %>% 
  mutate(scenario = "Memory Only", 
         rumination = if_else(rumination < 3.2, "-1 SD", 
                              if_else(rumination > 4, "+1 SD", 
                              "Mean"))) %>% 
  select(scenario, everything(), -Timepoint, -baseline)


bind_rows(best_contrast, worst_contrast, contagion_contrast, memory_contrast) %>% 
  select(-df, -statistic) %>% 
  mutate(p.value = round(p.value, 4), 
         significant = if_else(p.value < .05, "Sig", "NonSig"))

```
